\clearpage
\chapter*{Acknowledgements}
\label{ch:acknowledgements}
\begin{fullwidth}
  \noindent Thanks to Tod Machover for welcoming me into the Opera of
  the Future, for unending support, encouragement, and mentorship; for
  sharing your process, and for support integrating \thesis into
  \textit{Of Experience}.

\vspace{5mm}
\noindent Thanks to my readers James Andy Moorer and Joe Paradiso for
kind, and expert input, support, and guidance.

\vspace{5mm}
\noindent Thanks to Professor Alex Case for inspiring my love
for music, audio engineering, and illuminating the magical subtleties
of dynamic range compression, and for (only slightly begrudgingly)
writing me 1000 recommendations letters. No one is better suited to be
President of the Audio Engineering Society.

\vspace{5mm}
\noindent Thanks to Wonshik Choi and Niyom Lue for your infinite
patience, guidance, and for welcoming me to MIT in 2008. Only you
could have taught a music major to enjoy linux, DSP, and spectroscopy.

\vspace{5mm}
\noindent Thanks to Markus Phillips and Shawn Drost for directly and indirectly
giving me confidence as a software developer.

\vspace{5mm}
\noindent Thanks to my UMass Lowell Piano teachers for taking chance
with me, and putting up with me for four years. Anthony Mele,
Elizabeth Skavish, Bonnie Anderson, and Thomas Stumpf - You believed
in me before I did. I'm probably the only student ever who was lucky
enough to study with all four of you.

\vspace{5mm}
\noindent  Thanks to Gene Atwood for being considerate of everyone, and showing
me how important that is\ldots{} And for screaming in to a microphone when
I needed some screams.

\vspace{5mm}
\noindent Thanks to my roommates at The Red House, and Brainerd
Contextual Cooperative for love and support and home. For feeding me
lots of meals, for perspective, and for tolerating my student
lifestyle.

\vspace{5mm}
\noindent Thanks to my team at the MIT Media Lab. Ben Bloomberg for
being my peer and my mentor at the same time. Thanks for inviting me
into the Opera group in 2008 and again in 2014, teaching me a
tremendous amount about audio, and using your live sound prowess to
make \textit{De L'Exp\'{e}rience} sound amazing live. Bryn Bliska, for
conceiving Tempo Toy, which I only slightly stole and turned into
stochastic polytempic modulation, for letting me work out the maths,
and for humility, warmth and musical brilliance. Thanks to Rebecca
Kleinberger for infinite support and insight into everything from math
to hedgehogs to humans. David for peace, intensity, and robots. Akito
for your inspiring attitude, wisdom, projects, and humility. Kelly for
unending support, consideration, and sense of humor. Simone
for taking the ``im'' out of impossible (over and over and over
again).

\vspace{5mm}
\noindent Thanks to Helen Corless for being amazing supportive even
when I am in the absolute pits of grad student existence. Thank you
for always reminding me what music is really about, and for
challenging me like no one else can. Your kindness and wisdom make
the world a better place every day.

\vspace{5mm}
\noindent Thanks to my grandparents for leading by example, and teaching
kindness and dedication, and for endless support in education.

\vspace{5mm}
\noindent Thanks to my wonderful siblings: Hilary, Giles, and Felicity
for always inspiring me with kindness, honesty, and wisdom.

\vspace{5mm}
\noindent And thanks to my parents, Gwen and Mark for forcing me to get an
education before I was wise enough to know I wanted one. Thank you for
all your love and support and everything forever.
\end{fullwidth}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "CharlesHolbrow_MAS_Thesis"
%%% End:
